---
title: "PSTAT131 Homework 1"
author: "Shivani Kharva"
date: "2022-09-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 1  

## Machine Learning Main Ideas  

### Question 1  

From the textbook page 26, supervised learning is when, for each observation of the predictor measurements $x_{i}$, ${i}$ = 1,...,n, there is an associated response measurement $y_{i}$. In this case, the response is the supervisor. Models fitted to supervised learning tend to relate the response to the predictors. However, according to the textbook page 26, unsupervised learning is when, for each observation $i$,= 1,...,n, we observe a vector of measurements $x_{i}$, but no associated response $y_{i}$. We do not have a response variable that can supervise our analysis.  

Supervised and unsupervised learning are mainly different due to the existence of a response variable (the supervisor) for supervised learning and the nonexistence of a response variable (no supervisor) for unsupervised learning.  

### Question 2  

From the textbook page 28, regression models are those based on problems with a quantitative/numerical response whereas classification models are those based on problems involving a qualitative/categorical response (one with K different categories). Regression models are used for continuous outcomes, classification models are used for discrete outcomes.  

According to textbook page 29, these two models are both used in machine learning and often, the type of model to be used may be decided based on whether the response is qualitative (classification) or quantitative (regression). However, as mentioned on textbook page 29, whether the predictors are qualitative or quantitative is less important because many statistical methods can be applied regardless of the predictor variable type with proper coding for qualitative predictors in advance.  

### Question 3   HELP
(HELP what it mean by metric????)  

### Question 4   

**Descriptive models**: According to Lecture: Course Overview and Introduction slide 39, descriptive models visually illustrate trends in data. For example, using a line on a scatterplot is a descriptive model.

**Inferential models**: From Lecture: Course Overview and Introduction slide 39, inferential models aim to test theories. We are asking what features of the data are significant and attempting to learn whether there are any causal claims we can make. Inferential models aim to state a relationship between the outcome and the predictor(s).  

**Predictive models**: Based on Lecture: Course Overview and Introduction slide 39, predictive models aim to predict Y (the response) with minimum reducible error. Predictive models aim to determine what combination of features (predictors) fits the best. These models are not focusing on hypothesis tests.  

### Question 5  HELP

Mechanistic (aka parametric) predictive models assume that there are some parameters involved. According to Lecture: Course Overview and Introduction slide 38, mechanistic predictive models assume a parametric form for f, such as assuming f is a linear function. However, this means that this type of predictive model almost never actually matched the true unknown f. According to Lecture: Course Overview and Introduction slide 38, empirically-driven (non-parametric) predictive models do not make any assumptions about f and do not assume any parameters. Because of this, these type of predictive models require a larger number of observations.  

Based on Lecture: Course Overview and Introduction slide 38, The two model types are different because the mechanistic type assumes parameters and a parametric form of f whereas the empirically-driven type does not make any assumptions about parameters or f. Also, mechanistic model types require fewer observations than empirically-driven model types. Because of this difference, mechanistic model types require more parameters to be more flexible, while empirically-driven model types are much more flexible by default. However, according to Lecture: Course Overview and Introduction slide 38, the two model types can be similar in that these two model types may both lead to overfitting. In a mechanistic type model, if there are too many parameters, the model may become so flexible that it also fits random noise, which is overfitting. In an empirically-drive type model, since the model type is already much more flexible by default, it is more susceptible to overfitting by default as well.    

*(HELP IS THIS JUST OPINION???)*Mechanistic model is easier to understand because assumptions can be made about the model whereas in empirically-driven models, everything must be deduced from the data. 

*(HELP what is bias-variance tradeoff)*  
The bias-variance tradeoff is related to the use of mechanistic and empirically driven models in that the model type that you use may affect the bias-variance tradeoff. If a more empirically-driven model is used, such as drawing a curve that passes through every training observation, the method would have extremely low bias but high variance according to textbook page 36. If fitted to observations outside of the training set, the empirically-driven model would not be as accurate (because of overfitting). However, if a more mechanistic model is used, such as fitting a horizontal line to the data, the method might have very low variance but high bias according to textbook page 36. Using a more mechanistic model would allow for the training observations to be less fit to the chosen model, but using observations outside of the training set would not make the mechanistic model decrease in accuracy as with the empirically-driven model.  

### Question 6  HELP 

"Given a voter’s profile/data, how likely is it that they will vote in favor of the candidate?"  
This question is predictive because the campaign is attempting to predict the likelihood of the voter voting in favor of the candidate based on the features (predictors) presented in their profile/data. The aim is to predict Y, the likelihood that the voter will vote in favor of the candidate, with minimum reducible error (which is what a predictive model is from Lecture: Course Overview and Introduction slide 39).  

"How would a voter’s likelihood of support for the candidate change if they had personal contact with the candidate?"  
This question is inferential because the campaign is attempting to see how changing a specific feature related to the voter would affect their likelihood to support the candidate. Rather than using all features to best predict the voter's likelihood of supporting the candidate, the campaign is questioning how manipulating this one feature may affect the outcome. The aim is to test the relationship between the outcome (likelihood to vote in support of the voter) and a predictor (whether they had personal contact with the candidate) to see if any possible causal claims can be made (which is what an inferential model is from Lecture: Course Overview and Introduction slide 39).  

## Exploratory Data Analysis  
**Loading the tidyverse and other packages:**  
```{r}
library(tidyverse)
library(ISLR)
library(ggplot2)
library(corrplot)
data(mpg)
```

### Exercise 1
```{r}
hwy_hist <- ggplot(mpg, aes(hwy)) +
  geom_histogram()
  
hwy_hist
```
**Describe:**  
The histogram of highway miles per gallon is generally right skewed and bimodal. The two peaks are around 17 mpg and 27 mpg.  

### Exercise 2  (HELP)
```{r}
hwy_cty_scatter <- ggplot(mpg, aes(x=hwy, y=cty)) +
    geom_point()

hwy_cty_scatter
```
**Describe:**  
There is a strong positive relationship between hwy and cty. This means that, as the highway miles per gallon increases, the city miles per gallon also increases. Intuitively, this makes sense because those who get more mpg in any given city will also get more mpg on the highway.  

### Exercise 3  HELP (descending order??)
```{r}
manufacturer_bar <- ggplot(mpg, aes(y=manufacturer)) +
  geom_bar()

#NEED TO USE REORDER() WITHIN AES()? --> HOWWW

manufacturer_bar
```

Dodge produced the most cars. Lincoln produced the least cars.  

### Exercise 4  HELP WITH EXPLANATION
```{r}
hwy_cyl_boxplot <- ggplot(mpg, aes(x=hwy, group=cyl, fill=cyl)) +
  geom_boxplot()

hwy_cyl_boxplot
```
There is an apparent pattern shown with the boxplots. The highway miles per gallon appear to generally decrease as the cylinder count increases. The first quartile and median of the highway miles per gallon appear to decrease as each cylinder count increases. However, for the third quartile, the highway miles per gallon decrease as each cylinder count increases except at the 5 cylinder count.  (DO I NEED TO SAY ALL THIS?? can i just stop at second sentence?)  

### Exercise 5  HELP(doing it right?? year??)
```{r}
head(mpg)
M <- cor(mpg[, !(names(mpg) %in% c("manufacturer", "model", "trans", "drv","fl", "class"))])
corrplot(M,  method="square", type="lower", diag=FALSE)
```
cyl (number of cylinders) & displ (engine displacement) and hwy (highway miles per gallon) & cty (city miles per gallon) appear to be negatively correlated. According to Wikipedia, engine displacement is the measure of cylinder volume swept by all of the pistons of a piston engine, without including the combustion chamber. So, it would make sense that, as cyl increases, displ increases because there is more cylinder volume with more cylinders. Also, it makes sense that hwy and cty are positively correlated because, if individuals in certain cities are getting more miles per gallon on the highway, those same individuals are getting more miles per gallon in their respective cities.  

cty & displ, hwy & displ, cty & cyl, and hwy & cyl are all negatively correlated. According to Wikipedia, the lower the engine displacement, the less fuel an engine may consume. Therefore, it makes sense that cty & displ as well as hwy & displ are negatively correlated. Those with greater engine displacement consume more fuel, so they get fewer miles per gallon. This applies both on the highway and in their respective cities. So, as displ increases, hwy and cty naturally decrease. Also, for a similar reason, it makes sense that cty & cyl and hwy & cyl are negatively correlated because cyl & displ are positively correlated. So, a greater number of cylinders (which is associated with greater engine displacement) would lead to more fuel consumption. It makes sense that cty & cyl and hwy & cyl are negatively correlated because cars with a greater number of cylinders get fewer miles per gallon on the highway and in their respective cities.  

*Note*: I say "respective cities" here because, as mentioned before the cars that get high highway miles per gallon are also getting high city miles per gallon in their respective cities.  
